% first show that there exists an optimal seating where all ages are contiguous
% P = all people
% P' = P \ {those on the table with smallest}
% opt(P) = opt(P') + 1
\begin{solution}{Question 3.b}
    \begin{question}[]
        Suppose finally Alice invited $n_0$ out of her $n$ friends to the party. Her next task is to set a minimum number of dinner tables for her friends under the constraint that each table has a capacity of ten people and the age difference between members of each dining group should be at most ten years. Present a greedy algorithm to solve this problem in $O(n_0)$ time assuming the age of each person is an integer in the range $[10, 99]$.
    \end{question}
    \tcblower{}
    \begin{proof}[Solution]
        We have to implement an $O(n_0)$ greedy approach for the fore-mentioned question. Idea behind our algorithm is to create a frequency array which stores how many people of a certain age are there in the invited group from age $10$ to $100$. We then start grouping $10$ people together from the smallest to largest age. If the difference between youngest and oldest person exceeds $10$, we stop the group and create a new table. Following is the algorithm for the same:
        \begin{algorithm}[H]
            \caption{Find number of tables needed}
            \begin{algorithmic}[1]
                \Procedure{Seat}{$people$} \Comment{Finds the minimum number of tables needed}
                \State{$arr[100]\gets$initialised as zero array}\Comment{index 0 to 9 aren't used}
                \For{person in people} \Comment{get the frequency array}
                    \State{$arr[age(person)]\gets arr[age(person)]+1$}
                \EndFor{}
                \State{set $index, \ smallest, \ tcount\gets0$} \Comment{helpful intermediate variables}
                \State{set $toAdd \gets 10$}
                \While{$index<100$}
                    \If{$smallest=0$} \Comment{smallest=0 signifies no person on a table yet}
                        \State{$smallest \gets index$}
                    \EndIf{}
                    \If{$arr[index] = 0$} \Comment{no person of that age}
                        \State{$index\gets index+1$}
                    \ElsIf{$index>smallest+10$} \Comment{large age gap}
                        \State{$tcount \gets tcount+1$}
                        \State{$smallest \gets 0$}
                        \State{$toAdd \gets 10$}
                    \ElsIf{$arr[index]>toAdd$} \Comment{table occupancy can reach atmost $10$}
                        \State{$arr[index] \gets arr[index]-toAdd$}
                        \State{$tcount \gets tcount+1$}
                        \State{$toAdd \gets 10$}
                        \State{$smallest \gets 0$}
                    \Else{}
                        \State{$toAdd \gets toAdd-arr[index]$}
                        \State{$arr[index] \gets 0 $}
                        \State{$index \gets index+1$}
                    \EndIf{}
                \EndWhile{}
                \If{$toAdd \neq 10$} \Comment{take last non filled table into account}
                    \State{$tcount \gets tcount+1$}
                \EndIf{}
                \State{\Return{$tcount$}} \Comment{total number of tables required is returned}
                \EndProcedure{}
            \end{algorithmic}
        \end{algorithm}
        
        Now we shall prove the correctness of the algorithm:
        \subsection*{Initialisation}
        We generate a frequency array by declaring a zero array of size 100 and incrementing the value stored at an index by 1. The update at an index is done based on what age value the for loop comes across. Updated index is same as age. \\
        $index$ variable is used to refer the current index in the while loop.
        $toAdd$ variable shows how many more people are to be added at the recent table.
        $smallest$ denotes the smallest value of age of people sitting at a table. $smallest=0$ means that the table is empty.
        $tcount$ denotes the table number.
        All of the forementioned variables are initialsed to zero except toAdd which is initialised at 10.\\
        A while loop is run. The termination condition depends on the $index$ variable.
        The loop keeps incrementing index until it finds the index at which arr[index] is not 0.
        \subsection*{Maintenance}
        For maintenance, we look at the iterations of the while loop.\\
        \textbf{smallest = 0:}\\
        $smallest = 0$ implies that the table is new and there are no people seated on it.\\\\
        The following cases arise in the if-else block:\\
        \textbf{Case 1:} arr[index] is 0\\
        Here, we move to the next index since there are no people of this age invited to the party.\\
        \textbf{Case 2:} age gap between youngest person at table and index is $>10$\\
        Here, we cannot add the people at this index to the table since it violates our condition of the maximum age gap at table to be 10. We increase $tcount$ by 1 and reset values of $smallest$ and $toAdd$ to 0 and 10 respectively.\\
        \textbf{Case 3:} number of people with age same as index are more than $toAdd$\\
        In this case, the current table can be filled completely. We subtract the number of people to be added from arr[index] and increment $tcount$ by 1. $toAdd$ and $smallest$ are reinitialised.\\
        \textbf{Case 4:} number of people with age same as index are $\leq toAdd$\\
        Number of people of age same as index is not enough to fill the current table. We still allot them the table and move to next index. $toAdd$ is decremented by $arr[index]$. $arr[index]$ is set to 0.\\
        Since last group of statements is an else and no other conditions remain, our maintenance step is complete.

        \subsection*{Termination and Time Complexity}
        % \begin{enumerate}
            % \item According to the cases mentioned in the previous part, index increases by one when either $arr[index]=0$ or all people of same age are seated i.e in case 1 or case 5.
            % \item If case 2 occurs, $smallest$ is set to 0 and compulsorily leads to case 3 in next iteration.
            % \item If case 4 occurs, similar to case to, $smallest$ is set to 0 and leads to case 3 in next iteration.
            % \item case 3 allots the value of index which has atleast one person of that age to smallest who will be seated on the new table
            % \item It causes either of the case 1, 2, 4, 5 to repeat again.
        % \end{enumerate}
        In each iteration of the while loop, the value of $index$ increases by $1$ or the total sum of $arr$ decreases by some non-zero number or $tcount$ increments by $1$. Thus, the maximum number of iterations of the loop is bounded by:
        \begin{equation}
            \begin{split}
                100+sum(arr)+\max(tables)&=100+O(n_0)+O(n_0)\\
                &=O(n_0)
            \end{split}
        \end{equation}
        Therefore, the while loop runs for $O(n_0)$ iterations and each iteration takes $O(1)$ time since it involved making finite checks and updates all of which take $O(1)$ time.\\
        The loop from lines $3\text{--}5$ takes $O(n_0)$ time since it involves iterating over the entire array of size $n_0$ once.\\
        Therefore, the total time complexity of the algorithm is $O(n_0)$.

        \subsection*{Proof of Correctness}
        To prove the correctness of the algorithm, we first prove the following claim:
        \begin{claim}\label{claim:seatswap}
            If there exists an optimal seating where $p_i, p_k$ are on one table and $p_j$ is on another table such that $age(p_i)<age(p_j)<age(p_k)$, there exists another optimal seating where $p_i, p_j$ are on one table and $p_k$ is on another table.
        \end{claim}
        \begin{proof}
            Since $p_i$ and $p_k$ are on the same table, we know that $age(p_k)\leq age(p_{l1})+10$, where $p_{l1}$ is the person with least age on the table. Consider the person with smallest age on the table containing $p_j$, say $p_{l2}$. We know that $age(p_j)\leq age(p_{l2})+10$. Therefore, we have that:
            \begin{equation}
                age(p_k)\leq age(p_{l2})+10\text{, and }age(p_j)\leq age(p_{l1})+10
            \end{equation}
            Therefore, it is possible to swap the positions of $p_k$ and $p_j$. Thus, we have obtained an arrangement where $p_i, p_j$ are on the same table and $p_k$ is on a different table. This completes the proof of the claim.
        \end{proof}
        Consider the person with the smallest age, say $p_0$. Let the set of people be $P$ and the set of people arranged on the same table as $p_0$ using our algorithm be $A_0$. Let the optimal number of tables needed for a set of people be denoted by $opt(.)$. Now, we will show that,
        \begin{claim}\label{claim:optseat}
            $opt(P)=opt(P\setminus A_0)+1$
        \end{claim}
        \begin{proof}
            We first note that from Claim~\ref{claim:seatswap}, it is possible to generate a seating where no more swaps are possible. The set $A_0$ which is generated by our algorithm is one such arrangement for a single table. Thus, there exists an optimal arrangement with $A_0$ as one table. We will now prove the claim in two parts:
            \begin{enumerate}
                \item{$opt(P)\leq opt(P\setminus A_0)+1$}\\
                    An arrangement such that one table consists of people in $A_0$ and the remaining are arranged optimally will have the number of tables equal $opt(P\setminus A_0)+1$. This is by definition $\geq$ the most optimal seating for the set $P$ and thus this equality trivially holds.
                \item{$opt(P\setminus A_0)\leq opt(P)-1$}\\
                    We know that there exists an optimal arrangement which contains $A_0$ as one table. Therefore, the optimal arrangement for the remaining set of people, i.e., $P\setminus A_0$ can be done using atleast $opt(P)-1$ tables. Thus, we get that $opt(P\setminus A_0)\leq opt(P)-1$ which proves the second part of the claim.
            \end{enumerate}
            From $1$ and $2$, we get that $opt(P)=opt(P\setminus A_0)+1$, and this completes the proof of the claim.
        \end{proof}
        Therefore, from Claim~\ref{claim:optseat}, we have shown that our algorithm \textit{greedily} selects the most optimal seating. This completes the proof of correctness of our algorithm.
    \end{proof}
\end{solution}
